Dispersion Distance and the Matter Distribution of the Universe in Dispersion Space

### 2015-08-14 writeup/dm_space_distortions.tex
Copy edits.

Dispersion is readily measured from fast radio bursts (FRBs),
a new class of millisecond duration radio transients first reported by
\citet{2007Sci...318..777L}, which are the archetype of a standard ping.
Ten such bursts have been reported to date
\citep{2012MNRAS.425L..71K,2013Sci...341...53T, 2014ApJ...790..101S,
2014ApJ...792...19B, 2015MNRAS.447..246P,2015ApJ...799L...5R}, and the total
event rate has been estimated to be $3.3^{+5}_{-2.5} \times 10^3$ per sky per
day \cite{2015arXiv150500834R}.
Dispersion is characterized by a line-of-site
integral of the free electron density known the dispersion measure (DM). FRB
events are observed to have dispersion measures of order $\sim\!\!1000\,{\rm pc}
/ {\rm cm}^3$, greatly in excess of the galactic expectation of order
$\sim\!\!100\,{\rm pc} / {\rm cm}^3$ (depending on galactic latitude)
\cite{2014arXiv1412.4829D}.  They thus appear to be of extragalactic origin,
with their dispersion produced either by the intergalactic medium (IGM),
associated with structure along line of sight \citep{2013Sci...341...53T},
or the environment of the
source \citep{2014ApJ...785L..26L,2015arXiv150101341P,2015arXiv150505535C}.  If
the dispersion is dominated by electrons along the line of sight, then the
average dispersion--distance relation can be modeled
\citep{2014ApJ...797...71Z,2004MNRAS.348..999I, 2003ApJ...598L..79I} and the
sources appear to be at cosmological distances of order gigaparsecs.
Such cosmological distances permit the study of the clustering of FRB sources in
three dimensions, with an expectation that these events trace the large-scale
structure of the Universe on linear scales. We note that while our discussion
here is framed in terms of FRBs, other bright radio transients that fluctuate
on time scales shorter than the typical differential lag of $\sim 1s/100 \,{\rm
MHz} $ might also be useful as standard pings.

@@

In first measurements of the angular power spectra, uncertainty will be
dominated by the finite number of observed events, or shot noise.
The noise power spectrum on shells at a distance $\chi_i$ with width
$\Delta \chi$ is
\begin{equation}
    C^N_{\ell, ij} = \frac{\delta_{ij}}{\bar n_f(\chi_i)\chi_i^2\Delta\chi}.
\end{equation}
Here, to indicate that we are now working with quantities binned in the radial
direction, we have switched to $C_{\ell,ij}$ instead of
$C_\ell(\chi, \chi')$. The uncertainty on the angular
cross-power spectrum is then
\begin{align}
    (\Delta{C^{ss}_{\ell,ij}})^2 = 
    {\frac{1}{\ell(\ell + 1)\Delta \ell f_{\rm sky}}}
        \left[C^N_{\ell,ii} C^N_{\ell,jj} +  (C^N_{\ell,ij})^2\right],
\end{align}
where the second term in brackets only contributes for $i = j$.

>>>>

Dispersion is readily measured from fast radio bursts (FRBs),
a new class of millisecond duration radio transients first reported by
\citet{2007Sci...318..777L}, which are the archetype of a standard ping.
Ten such bursts have been reported to date
\citep{2012MNRAS.425L..71K,2013Sci...341...53T, 2014ApJ...790..101S,
2014ApJ...792...19B, 2015MNRAS.447..246P,2015ApJ...799L...5R}, and the total
event rate has been estimated to be $3.3^{+5}_{-2.5} \times 10^3$ per sky per
day \cite{2015arXiv150500834R}.
Dispersion is characterized by a line-of-site
integral of the free electron density known the dispersion measure (DM). FRB
events are observed to have dispersion measures of order $\sim\!\!1000\,{\rm pc}
/ {\rm cm}^3$, greatly in excess of the galactic expectation of order
$\sim\!\!100\,{\rm pc} / {\rm cm}^3$ (depending on galactic latitude)
\cite{2015MNRAS.451.4277D}.  They thus appear to be of extragalactic origin,
with their dispersion produced either by the intergalactic medium,
associated with structure along line of sight \citep{2013Sci...341...53T},
or the environment of the
source \citep{2014ApJ...785L..26L,2015ApJ...807..179P,2015arXiv150505535C}.  If
the dispersion is dominated by electrons along the line of sight, then the
average dispersion--distance relation can be modeled
\citep{2014ApJ...797...71Z,2004MNRAS.348..999I, 2003ApJ...598L..79I} and the
sources appear to be at cosmological distances of order gigaparsecs.
Such cosmological distances permit the study of the clustering of FRB sources in
three dimensions, with an expectation that these events trace the large-scale
structure of the Universe on linear scales. We note that while our discussion
here is framed in terms of FRBs, other bright radio transients that fluctuate
on time scales shorter than the typical differential lag of $\sim 1s/100 \,{\rm
MHz} $ might also be useful as standard pings.

@@

In first measurements of the angular power spectra, uncertainty will be
dominated by the finite number of observed events, or shot noise.
The noise power spectrum on shells at a distance $\chi_i$ and with width
$\Delta \chi$ is
\begin{equation}
    C^N_{\ell, ij} = \frac{\delta_{ij}}{\bar n_f(\chi_i)\chi_i^2\Delta\chi}.
\end{equation}
Here, to indicate that we are now working with quantities binned in the radial
direction, we have switched to $C_{\ell,ij}$ instead of
$C_\ell(\chi, \chi')$. The uncertainty on the angular
cross-power spectrum is then
\begin{align}
    (\Delta{C^{ss}_{\ell,ij}})^2 = 
    {\frac{1}{\ell(\ell + 1)\Delta \ell f_{\rm sky}}}
        \left[C^N_{\ell,ii} C^N_{\ell,jj} +  (C^N_{\ell,ij})^2\right],
\end{align}
where the second term in brackets only contributes for $i = j$.
